---
title: "AQUASTAT World Water Stress"
subtitle: ""
author: "Evan Rank"
institute: ""
date: "2021/05/09 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```
# Dataset

AQUASTAT is a publicly available, free information system on water resources and agriculture management.

It is maintained by the Food and Agriculture Organization of the United Nations.

It can be accessed at http://www.fao.org/aquastat/en/. Click on "Databases" to access the system (a short survey may be prompted).

I will provide the CSV of my own query, but you can make your own to access different data.
---

# Libraries and Data

```{r}
library(here)
library(tidyverse)
library(maps)
library(mapdata)
library(mapproj)
library(ggrepel)

aquastat <- read_csv(here("Data", "aquastat_flat.csv"))
```

---

# Goal of Inquiry

* What countries are experiencing high water stress conditions today?
  + Water stress occurs when the demand for water exceeds available water, whether the restriction is by amount or quality.
  
* What is the recent history of water stress? Any trends?

* Hypothesis: Economic inequality makes it difficult to correct water availability issues.

---
# Cleaning Data

```{r}
aquastat_wide <- aquastat %>% 
  select(Area, Variable_Name, Year, Value) %>% 
  filter(complete.cases(.)) %>% 
  pivot_wider(names_from = Variable_Name, values_from = Value)

colClean <- function(x){ colnames(x) <- gsub(" ", "_", colnames(x)); x }

aquastat_clean <- colClean(aquastat_wide)
```

---
# Recent Water Stress Data - Mapping
* Prep to Join with "world"
```{r}
world <- map_data("world")

aquastat_stress <- aquastat_clean %>% 
  select(Area, SDG_6.4.2._Water_Stress, Year) %>% 
  filter(complete.cases(.)) %>% 
  rename(region = Area)

#Check for unjoined regions
anti_2017 <- aquastat_stress %>% 
  filter(Year==2017) %>% 
  anti_join(world)

head(anti_2017)
```
---
# Rename to Join

```{r}
aquastat_stress1 <- aquastat_clean %>% 
  select(Area, SDG_6.4.2._Water_Stress, Year) %>% 
  filter(complete.cases(.)) %>% 
  rename(region = Area) %>% 
  mutate(region = str_replace(region, "United States of America", "USA")) %>% 
  mutate(region = str_replace(region, "Russian Federation", "Russia")) %>% 
  mutate(region = str_replace(region, "Democratic People's Republic of Korea", "North Korea")) %>% 
  mutate(region = str_replace(region, "Venezuela \\(Bolivarian Republic of\\)", "Venezuela")) %>% 
  mutate(region = str_replace(region, "Viet Nam", "Vietnam")) %>%
  mutate(region = str_replace(region, "Bolivia \\(Plurinational State of\\)", "Bolivia")) %>%
  mutate(region = str_replace(region, "Brunei Darussalam", "Brunei")) %>%
  mutate(region = str_replace(region, "Cabo Verde", "Cape Verde")) %>%
  mutate(region = str_replace(region, "Côte d'Ivoire", "Ivory Coast")) %>%
  mutate(region = str_replace(region, "Iran \\(Islamic Republic of\\)", "Iran")) %>%
  mutate(region = str_replace(region, "Republic of Korea", "South Korea")) %>%
  mutate(region = str_replace(region, "Republic of Moldova", "Moldova")) %>%
  mutate(region = str_replace(region, "United Kingdom", "UK")) %>%
  mutate(region = str_replace(region, "United Republic of Tanzania", "Tanzania")) %>%
  mutate(region = str_replace(region, "Syrian Arab Republic", "Syria")) %>% 
  mutate(region = str_replace(region, "Congo", "Republic of Congo")) %>% 
  mutate(region = str_replace(region, "Democratic Republic of the Republic of Congo", "Democratic Republic of the Congo"))
```
---
# Join with "world"
```{r}
stress_2017 <- aquastat_stress1 %>% 
  filter(Year==2017) %>% 
  left_join(world)
```
---

# Map Data

.pull-left[
```{r, stress_2017, eval=FALSE, warning=FALSE, message=FALSE}
ggplot()+
  geom_polygon(data=stress_2017, aes(x=long, y=lat, group=group, fill=SDG_6.4.2._Water_Stress), color="black")+
  theme_minimal()+
  coord_map("mercator", xlim=c(-180, 180))+
  scale_fill_gradient(trans="log10", low="#B50A2AFF", high="#0E84B4FF")+
  labs(title="World Water Stress, ~2017",
       x="",
       y="",
       fill="Water Stress (log10)")+
  theme(plot.title = element_text(hjust = 0.5, color = "#333544FF"),
        legend.title = element_text(color = "#333544FF"),
        panel.background= element_rect(color="grey12"))


```
]
.pull-right[
```{r stress_2017-out, ref.label="stress_2017", echo=FALSE}

```
]

---
```{r stress_2017-out-1, ref.label="stress_2017", echo=FALSE}

```

---
# Deciding what to fit in a time-series

```{r}
aquastat_stress2 <- aquastat_clean %>% 
  select(Area, SDG_6.4.2._Water_Stress, Year, Total_population) %>% 
  filter(complete.cases(.)) %>% 
  filter(Year==1992)

```
---
# Subset selected areas
```{r}
aquastat_high_pop <- aquastat_clean %>% 
  select(Area, SDG_6.4.2._Water_Stress, Year, Total_population) %>% 
  filter(complete.cases(.)) %>% 
  filter(str_detect(Area, "China|India|United States of America|Indonesia|Brazil|Russian Federation|Japan|Pakistan|Nigeria|Mexico|Germany|Viet Nam|Iran \\(Islamic Republic of\\)|Egypt|United Kingdom"))
```
---
# Line chart in time series

.pull-left[
```{r, stress_by_year, eval=FALSE, warning=FALSE, message=FALSE}
ggplot(aquastat_high_pop, aes(x=Year, y= SDG_6.4.2._Water_Stress, color=Area, label=Area))+
  geom_line(size=1.5)+
  geom_text_repel(data = aquastat_high_pop %>% filter(Year == 2017), nudge_x = 2, size=4, show.legend = FALSE)+
  labs(x= "Year",
       y= "Water Stress",
       title= "Water Stress by Year",
       subtitle="In Worldwide High-Population Nations",
       color= "Country")+
  scale_color_viridis_d()+
  theme(plot.title=element_text(hjust=0.5),
        plot.subtitle=element_text(hjust=0.5))
```
]

.pull-right[
```{r stress_by_year-out, ref.label="stress_by_year", echo=FALSE}

```
]
---
# Is this related to economic success?

* As before but selecting with GDP
```{r}
aquastat_WS_GDP <- aquastat_clean %>% 
  select(Area, SDG_6.4.2._Water_Stress, Year, Total_population, `Gross_Domestic_Product_(GDP)`) %>% 
  filter(complete.cases(.)) %>% 
  filter(str_detect(Area, "China|India|United States of America|Indonesia|Brazil|Russian Federation|Japan|Pakistan|Nigeria|Mexico|Germany|Viet Nam|Iran \\(Islamic Republic of\\)|Egypt|United Kingdom"))
```

---

# Relationship between GDP and changes in water stress
.pull-left[
```{r, stress_gdp_year, eval=FALSE, warning=FALSE, message=FALSE}
  ggplot(aquastat_WS_GDP, aes(x=SDG_6.4.2._Water_Stress, y=`Gross_Domestic_Product_(GDP)`, color=Area))+ 
    geom_point(size=4)+
    facet_wrap(~Year)+
    labs(x="Water Stress",
         y= "Gross Domestic Product",
         title="Water stress vs GDP Over Time",
         subtitle="In Worldwide High-Population Nations")+
    scale_color_viridis_d()
  
```
]
.pull-right[
```{r stress_gdp_year-out, ref.label="stress_gdp_year", echo=FALSE}

```
]
---

```{r stress_gdp_year-out-1, ref.label="stress_gdp_year", echo=FALSE}

```
---

# Thanks for a great semester, BIOL 551!

* Slides made with Xaringan package
* Data from http://www.fao.org/aquastat/en/

---









